---
title: "Process metabarcoding data"
format: html
editor: visual
execute: 
  eval: false
---

# Obtain 18S rRNA data from prior sequence run

R libraries

```{r}
library(tidyverse)
library(phyloseq)
library(decontam)
```

```{r}
merged_tax <- read_delim("../microeuks_deepbiosphere_datamine/microeuk-amplicon-survey/data-input/taxonomy.tsv", delim = "\t")

merged_asv <- read_delim("../microeuks_deepbiosphere_datamine/microeuk-amplicon-survey/data-input/microeuk-merged-asv-table.tsv", delim = "\t", skip = 1)

metadata <- read.delim("../microeuks_deepbiosphere_datamine/microeuk-amplicon-survey/data-input/samplelist-metadata.txt")
 
asv_wtax <- merged_asv %>%
  select(FeatureID = '#OTU ID', everything()) %>%
  pivot_longer(cols = !FeatureID,
               names_to = "SAMPLE", values_to = "value") %>%
  left_join(merged_tax, by = c("FeatureID" = "Feature ID")) %>%
  left_join(metadata) %>%
  filter(SITE == "GordaRidge" | SITE == "substrate" | SITE == "Laboratory") %>%
  filter(!grepl("Siders_", SAMPLE)) %>%
  filter(!(grepl("T0", SAMPLE))) %>%
  filter(!(grepl("T24", SAMPLE))) %>%
  filter(!(grepl("T36", SAMPLE))) %>%
  mutate(DATASET = case_when(
    grepl("_GR_", SAMPLE) ~ "GR",
    grepl("Gorda", SAMPLE) ~ "GR",
    grepl("_MCR_", SAMPLE) ~ "MCR",
    grepl("Axial", SAMPLE) ~ "Axial",
  TRUE ~ "Control or blank")) %>%
    separate(Taxon, c("Domain", "Supergroup",
                  "Phylum", "Class", "Order",
                  "Family", "Genus", "Species"), sep = ";", remove = FALSE)

# fix naming, some controls sequenced separately.
gr_substrate_fluid_asvs <- asv_wtax %>%
  mutate(SAMPLE_tmp = case_when(
    Sample_actual == "" ~ SAMPLE,
    TRUE ~ Sample_actual
  )) %>%
  select(-SAMPLE) %>%
  select(SAMPLE = SAMPLE_tmp, everything()) %>%
  filter(value > 0)

# save(gr_substrate_fluid_asvs, file = "input-data/gordaridge_18S_asvs.RData")
```

## Run decontam on 18S rRNA data

## Decontaminate sequence library

Import sample description text file, import as phyloseq library, and remove potential contaminate ASVs and sequences. Catalog total number of ASVs and sequences removed from analysis.

### Import as phyloseq objects

```{r Extract taxa and sample info as matrix for phyloseq input}
# head(gr_substrate_fluid_asvs)
tax_matrix <- gr_substrate_fluid_asvs %>% 
  select(FeatureID, Taxon) %>% 
  distinct() %>% 
  separate(Taxon, c("Domain", "Supergroup", 
                  "Phylum", "Class", "Order",
                  "Family", "Genus", "Species"), sep = ";", remove = FALSE) %>% 
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix

asv_matrix <- gr_substrate_fluid_asvs %>% 
  select(FeatureID, SAMPLE, value) %>% 
  pivot_wider(names_from = "SAMPLE", values_fill = 0, values_from = value) %>% 
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix

# Align row names for each matrix
rownames(tax_matrix) <- row.names(asv_matrix)

metadata_cones <- gr_substrate_fluid_asvs %>% 
  select(SAMPLE, Type, VENT, SITE, SAMPLETYPE, Sample_or_Control) %>% 
  distinct() %>% 
  column_to_rownames(var = "SAMPLE")
```

```{r Phyloseq import}
# Import asv and tax matrices
ASV = otu_table(asv_matrix, taxa_are_rows = TRUE)
TAX = tax_table(tax_matrix)
phylo_obj <- phyloseq(ASV, TAX)

# Import metadata as sample data in phyloseq
samplenames <- sample_data(metadata_cones)

# join as phyloseq object
physeq_wnames = merge_phyloseq(phylo_obj, samplenames)
```

### Identify contaminant ASVs

In addition to shipboard milliQ blank samples, each substrate type had a 'blank' control, which was sampled at the same time, but never deployed in the microcolonizers (only processed at the same time in the lab).

```{r Assign control samples}
# When "Control" appears in "Sample_or_Control column, this is a negative control"
sample_data(physeq_wnames)$is.neg <- sample_data(physeq_wnames)$Sample_or_Control == "Control"
```

```{r Identify contaminant ASVs}
# ID contaminants using Prevalence information
contam_prev <- isContaminant(physeq_wnames, 
                               method="prevalence", 
                               neg="is.neg", 
                               threshold = 0.5, normalize = TRUE) 

# Report number of ASVs IDed as contamintants
table(contam_prev$contaminant)
```

> 0.5 - this threshold will ID contaminants in all samples that are more prevalent in negative controls than in positive samples.

### Remove problematic ASVs

```{r Assign contaminant ASVs}
# Subset contaminant ASVs
contams <- filter(contam_prev, contaminant == "TRUE")
list_of_contam_asvs <- as.character(row.names(contams))
# length(list_of_contam_asvs)

taxa_contam <- as.data.frame(tax_matrix) %>% 
  rownames_to_column(var = "FeatureID") %>% 
  filter(FeatureID %in% list_of_contam_asvs)
# head(taxa_contam)
```

```{r Compare decontaminated data with original}
# View(asv_wtax)
asv_wtax_decon <- gr_substrate_fluid_asvs %>% 
  filter(!(FeatureID %in% list_of_contam_asvs)) %>% 
  filter(!(Sample_or_Control == "Control"))

tmp_orig <- (gr_substrate_fluid_asvs %>% filter(!(Sample_or_Control == "Control")))

# Stats on lost
x <- length(unique(tmp_orig$FeatureID)); x
y <- length(unique(asv_wtax_decon$FeatureID)); y
100*((y-x)/x) #0.43% of ASVs lost
a <- sum(tmp_orig$value);a #3.1 million
b <- sum(asv_wtax_decon$value);b #2.89 million 
100*((b-a)/a)
# Lost 5.5% of sequences from whole dataset.

## Subsample to clean ASVs
asv_wtax_wstats <- gr_substrate_fluid_asvs %>% 
  mutate(DECONTAM = case_when(
    FeatureID %in% list_of_contam_asvs ~ "FAIL",
    TRUE ~ "PASS"
  ))
```

## QC 18S rRNA gene dataset

```{r}
head(asv_wtax_wstats)

asv_wtax_18 <- asv_wtax_wstats |> 
  filter(Sample_or_Control == "Sample") |> 
  filter(DECONTAM == "PASS") |> 
  select(SAMPLE, FeatureID, SEQUENCE_COUNT = value, Taxon, Domain,
         Supergroup, Phylum, Class, Order, Family, Genus, Species, VENT, COORDINATES, SITE, SAMPLEID, SAMPLETYPE, DEPTH, temp, pH, percseawater, mg, h2, h2s, ch4, ProkConc, MC = Type) |> 
  add_column(dataset = "18S") |> 
  mutate(Substrate = case_when(
    SAMPLETYPE == "Microcolonizer" ~ VENT
  ))
```

# Obtain 16S rRNA data frames

```{r}
df_16 <- read_delim("input-data/samples16s-asv-table.tsv", delim = "\t", skip = 1)
head(df_16)
tax_16 <- read_delim("input-data/taxonomy.tsv", delim = "\t")
head(tax_16)
```

```{r}
metadata_16s <- read.csv("input-data/manifest-16s.csv")
# head(metadata_16s)
```

```{r}
head(tax_16)
```

```{r}
asv_wtax_16 <- df_16 |> 
  select(`Feature ID` = `#OTU ID`, everything()) |> 
  left_join(tax_16) |> 
  pivot_longer(cols = starts_with("MC"), names_to = "SAMPLES", values_to = "SEQUENCE_COUNT") |> 
  select(FEATUREID = `Feature ID`, Taxon, SAMPLES, SEQUENCE_COUNT) |> 
  mutate(TAXON = str_replace_all(Taxon, ".__", "")) |> 
  separate(TAXON, into = c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species"), remove = FALSE, sep = "; ") |> 
  left_join((metadata_16s |> select(SAMPLES = `sample.id`, MC, Substrate))) |> 
  mutate(MC = paste("MC", MC, sep = "")) |> 
  add_column(dataset = "16S")

# head(asv_wtax_16)
```

```{r}
save(asv_wtax_16, asv_wtax_18, file = "input-data/GR_Microcolonizer_data.RData")
```
