---
title: "Gorda Ridge Cones"
author: "Sarah Hu"
date: "8/19/2021"
output: 
  html_document:
    number_sections: true
    theme: cosmo
    highlight: kate
    collapsed: false
    toc: true
    toc_depth: 4
    toc_float: true

knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Library loading initial}
library(tidyverse); library(cowplot)
library(phyloseq); library(decontam)
library(plotly); library(lubridate)
```

# Merged data from QIIME2 analysis

Run below to repeat ASV compilation from raw QIIME2 output.
```{r Join raw taxonomy and ASV data}
# merged_tax <- read_delim("/Users/sarahhu/Desktop/Projects/microeuks_deepbiosphere_datamine/microeuk-amplicon-survey/data-input/taxonomy.tsv", delim = "\t")
# merged_asv <- read_delim("/Users/sarahhu/Desktop/Projects/microeuks_deepbiosphere_datamine/microeuk-amplicon-survey/data-input/microeuk-merged-asv-table.tsv", delim = "\t", skip = 1)
# metadata <- read.delim("/Users/sarahhu/Desktop/Projects/microeuks_deepbiosphere_datamine/microeuk-amplicon-survey/data-input/samplelist-metadata.txt")
# 
# asv_wtax <- merged_asv %>%
#   select(FeatureID = '#OTU ID', everything()) %>%
#   pivot_longer(cols = !FeatureID,
#                names_to = "SAMPLE", values_to = "value") %>%
#   left_join(merged_tax, by = c("FeatureID" = "Feature ID")) %>%
#   left_join(metadata) %>%
#   filter(SITE == "GordaRidge" | SITE == "substrate" | SITE == "Laboratory") %>%
#   filter(!grepl("Siders_", SAMPLE)) %>%
#   filter(!(grepl("T0", SAMPLE))) %>%
#   filter(!(grepl("T24", SAMPLE))) %>%
#   filter(!(grepl("T36", SAMPLE))) %>%
#   mutate(DATASET = case_when(
#     grepl("_GR_", SAMPLE) ~ "GR",
#     grepl("Gorda", SAMPLE) ~ "GR",
#     grepl("_MCR_", SAMPLE) ~ "MCR",
#     grepl("Axial", SAMPLE) ~ "Axial",
#   TRUE ~ "Control or blank")) %>%
#     separate(Taxon, c("Domain", "Supergroup",
#                   "Phylum", "Class", "Order",
#                   "Family", "Genus", "Species"), sep = ";", remove = FALSE)
# 
# # fix naming, some controls sequenced separately.
# gr_substrate_fluid_asvs <- asv_wtax %>%
#   mutate(SAMPLE_tmp = case_when(
#     Sample_actual == "" ~ SAMPLE,
#     TRUE ~ Sample_actual
#   )) %>%
#   select(-SAMPLE) %>%
#   select(SAMPLE = SAMPLE_tmp, everything()) %>%
#   filter(value > 0)

# View(gr_substrate_fluid_asvs)
# View(gr_substrate_fluid_asvs %>% filter(Sample_or_Control == "Control") %>% select(SAMPLE) %>% distinct())
# save(gr_substrate_fluid_asvs, file = "/Users/sarahhu/Desktop/Projects/GordaRidge-microcolonizers/microcolonizers-GordaRidge/input-data/asv-table.RData")
```

# Import microcolonizer data
```{r}
load("input-data/asv-table.RData", verbose = TRUE)
```

### Quick look - unprocessed data
Barplots to show total number of sequences and total number of ASVs
```{r Plot total ASV and sequences, fig.height=12, fig.width=19}
plot_grid(gr_substrate_fluid_asvs %>% 
  filter(value > 0) %>% 
  ggplot(aes(x = SAMPLE)) +
  geom_bar(stat = "count", width = 0.9) +
  labs(y = "Total ASVs") +
  coord_flip() +
  theme_linedraw() +
  facet_grid(DATASET ~ ., scale = "free", space = "free") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1),
        strip.background = element_blank(), strip.text = element_text(color = "black")),
  gr_substrate_fluid_asvs %>% 
  group_by(SAMPLE, SITE, Domain, DATASET) %>% 
  summarise(SUM_SEQ_DOMAIN = sum(value)) %>% 
  ggplot(aes(x = SAMPLE, y = SUM_SEQ_DOMAIN, fill = Domain)) +
  geom_bar(stat = "identity", color = "black", width = 0.9) +
  labs(y = "Total Sequences") +
  coord_flip() +
  theme_linedraw() +
  facet_grid(DATASET ~ ., scale = "free", space = "free") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1),
        strip.background = element_blank(), strip.text = element_text(color = "black"),
        legend.position = "bottom"),
  ncol = 2)
```

### Table reporting total ASVs and sequences
```{r Total ASVs raw}
# View(gr_substrate_fluid_asvs %>% filter(value > 0) %>% 
#        group_by(SAMPLE, DATASET, SITE) %>% 
#        summarise(SEQ_SUM = sum(value),
#                  ASV_COUNT = n()))
```


## Decontaminate sequence library
Import sample description text file, import as phyloseq library, and remove potential contaminate ASVs and sequences. Catalog total number of ASVs and sequences removed from analysis.

### Import as phyloseq objects
```{r Extract taxa and sample info as matrix for phyloseq input}
# head(gr_substrate_fluid_asvs)
tax_matrix <- gr_substrate_fluid_asvs %>% 
  select(FeatureID, Taxon) %>% 
  distinct() %>% 
  separate(Taxon, c("Domain", "Supergroup", 
                  "Phylum", "Class", "Order",
                  "Family", "Genus", "Species"), sep = ";", remove = FALSE) %>% 
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix

asv_matrix <- gr_substrate_fluid_asvs %>% 
  select(FeatureID, SAMPLE, value) %>% 
  pivot_wider(names_from = "SAMPLE", values_fill = 0, values_from = value) %>% 
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix

# Align row names for each matrix
rownames(tax_matrix) <- row.names(asv_matrix)

metadata_cones <- gr_substrate_fluid_asvs %>% 
  select(SAMPLE, Type, VENT, SITE, SAMPLETYPE, Sample_or_Control) %>% 
  distinct() %>% 
  column_to_rownames(var = "SAMPLE")
```

```{r Phyloseq import}
# Import asv and tax matrices
ASV = otu_table(asv_matrix, taxa_are_rows = TRUE)
TAX = tax_table(tax_matrix)
phylo_obj <- phyloseq(ASV, TAX)

# Import metadata as sample data in phyloseq
samplenames <- sample_data(metadata_cones)

# join as phyloseq object
physeq_wnames = merge_phyloseq(phylo_obj, samplenames)
# colnames(ASV)

## Check
# physeq_wnames
# View(gr_substrate_fluid_asvs %>% filter(value > 0))
```

### Identify contaminant ASVs

In addition to shipboard milliQ blank samples, each substrate type had a 'blank' control, which was sampled at the same time, but never deployed in the microcolonizers (only processed at the same time in the lab). 

```{r Assign control samples}
# When "Control" appears in "Sample_or_Control column, this is a negative control"
sample_data(physeq_wnames)$is.neg <- sample_data(physeq_wnames)$Sample_or_Control == "Control"
```

```{r Identify contaminant ASVs}
# ID contaminants using Prevalence information
contam_prev <- isContaminant(physeq_wnames, 
                               method="prevalence", 
                               neg="is.neg", 
                               threshold = 0.5, normalize = TRUE) 

# Report number of ASVs IDed as contamintants
table(contam_prev$contaminant)
```
> 0.5 - this threshold will ID contaminants in all samples that are more prevalent in negative controls than in positive samples.

### Remove problematic ASVs
```{r Assign contaminant ASVs}
# Subset contaminant ASVs
contams <- filter(contam_prev, contaminant == "TRUE")
list_of_contam_asvs <- as.character(row.names(contams))
# length(list_of_contam_asvs)

taxa_contam <- as.data.frame(tax_matrix) %>% 
  rownames_to_column(var = "FeatureID") %>% 
  filter(FeatureID %in% list_of_contam_asvs)
# head(taxa_contam)
```


```{r Compare decontaminated data with original}
# View(asv_wtax)
asv_wtax_decon <- gr_substrate_fluid_asvs %>% 
  filter(!(FeatureID %in% list_of_contam_asvs)) %>% 
  filter(!(Sample_or_Control == "Control"))

tmp_orig <- (gr_substrate_fluid_asvs %>% filter(!(Sample_or_Control == "Control")))

# Stats on lost
x <- length(unique(tmp_orig$FeatureID)); x
y <- length(unique(asv_wtax_decon$FeatureID)); y
100*((y-x)/x) #0.43% of ASVs lost
a <- sum(tmp_orig$value);a #3.1 million
b <- sum(asv_wtax_decon$value);b #2.89 million 
100*((b-a)/a)
# Lost 5.5% of sequences from whole dataset.

## Subsample to clean ASVs
asv_wtax_wstats <- gr_substrate_fluid_asvs %>% 
  mutate(DECONTAM = case_when(
    FeatureID %in% list_of_contam_asvs ~ "FAIL",
    TRUE ~ "PASS"
  ))
```

```{r, fig.height=12, fig.width=10}
plot_grid(asv_wtax_wstats %>% 
  filter(value > 0) %>% 
  ggplot(aes(x = SAMPLE, fill = DECONTAM)) +
  geom_bar(stat = "count", width = 0.9, color = "black") +
  labs(y = "Total ASVs") +
  coord_flip() +
  theme_linedraw() +
  facet_grid(DATASET ~ ., scale = "free", space = "free") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1),
        strip.background = element_blank(), strip.text = element_text(color = "black"),
        legend.position = "bottom"),
  asv_wtax_wstats %>% 
  group_by(SAMPLE, SITE, DECONTAM, DATASET) %>% 
  summarise(SUM_SEQ_DOMAIN = sum(value)) %>% 
  ggplot(aes(x = SAMPLE, y = SUM_SEQ_DOMAIN, fill = DECONTAM)) +
  geom_bar(stat = "identity", color = "black", width = 0.9) +
  labs(y = "Total Sequences") +
  coord_flip() +
  theme_linedraw() +
  facet_grid(DATASET ~ ., scale = "free", space = "free") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1),
        strip.background = element_blank(), strip.text = element_text(color = "black"),
        legend.position = "bottom"),
  ncol = 2)
```


# Subset Gorda Ridge Cones and _in situ_

Isolate Gorda Ridge substrate samples and the in situ samples from Mt Edwards, plume at Mt Edwards and the background seawater. Remove in situ
```{r Isolate in situ only}
# colnames(asv_wtax_wstats)
# unique(asv_wtax_wstats$Sample_actual)
# Isolate blank substrate for GR cones

asv_GR <- asv_wtax_wstats %>% 
      filter(Sample_or_Control != "Control") %>% # rm controls
      filter(SITE == "substrate" | VENT == "Mt Edwards" | 
                VENT == "Mt Edwards Plume" | VENT == "Deep seawater" | VENT == "Shallow seawater") %>% 
      filter(DATASET != "Axial", DATASET != "MCR") %>% 
      filter(!grepl("_expTf_", SAMPLE)) %>% 
      filter(value > 0) %>% 
      filter(DECONTAM == "PASS") #%>% 
      # bind_rows(asv_substrate_blanks)

# Get quick stats on totals
sum(asv_GR$value) # 2.3 million sequences
length(unique(asv_GR$FeatureID)) #3370 ASVs

# View(asv_GR)
# unique(asv_GR$VENT)
```


Additional sample QC, check replicates, and determine if replicates should be averaged.
```{r Investigate sequence and sample QC, fig.height=9, fig.width=14}
plot_grid(asv_GR %>% 
  group_by(SAMPLE, VENT, DATASET, SAMPLETYPE, Domain) %>% 
  summarise(seqsum_var = sum(value),
            asvcount_var = n()) %>% 
  pivot_longer(ends_with("_var"), names_to = "VARIABLE") %>% 
  ggplot(aes(x = SAMPLE, y = value, fill = Domain)) +
    geom_bar(color = "black", stat = "identity", position = "fill") +
    facet_grid(VARIABLE ~ SAMPLETYPE, space = "free", scales = "free") +
  scale_y_continuous(expand = c(0,0)) +
  theme_linedraw() +
  scale_fill_brewer(palette = "Paired") +
  theme(strip.background = element_blank(), strip.text = element_text(color = "black"),
        axis.text.x = element_text(color = "black", angle = 90, hjust = 1, vjust = 0.5),
        legend.position = "bottom"),
  asv_GR %>% 
  group_by(SAMPLE, VENT, DATASET, SAMPLETYPE, Domain) %>% 
  summarise(seqsum_var = sum(value),
            asvcount_var = n()) %>% 
  pivot_longer(ends_with("_var"), names_to = "VARIABLE") %>% 
  ggplot(aes(x = SAMPLE, y = value, fill = Domain)) +
    geom_bar(color = "black", stat = "identity", position = "stack") +
    facet_grid(VARIABLE ~ SAMPLETYPE, space = "free_x", scales = "free") +
  scale_y_continuous(expand = c(0,0)) +
  theme_linedraw() +
  scale_fill_brewer(palette = "Paired") +
  theme(strip.background = element_blank(), strip.text = element_text(color = "black"),
        axis.text.x = element_text(color = "black", angle = 90, hjust = 1, vjust = 0.5),
        legend.position = "bottom"),
ncol = 2)
```

Based on above plot, remove "BSW020" sterivex filter. 

## Taxonomy bar plots

Supergroup distribution by background fluid vs. what was isolated from substrates.

Factor supergroup names
```{r}
tmp <- asv_GR %>% 
  filter(Domain == "Eukaryota") %>%
  filter(Sample_or_Control != "Control") %>% 
  # unite(SampleIdentifier, VENT, SAMPLETYPE, sep = " ", remove = FALSE) %>% 
  mutate(Supergroup = ifelse(is.na(Supergroup), "Unknown Eukaryota", Supergroup),
         Phylum = ifelse(is.na(Phylum), "Unknown", Phylum),
         Phylum = ifelse(Phylum == "Alveolata_X", "Ellobiopsidae", Phylum),
         Supergroup = ifelse(Supergroup == "Alveolata", paste(Supergroup, Phylum, sep = "-"), Supergroup))
colors_tax <- c("#f1eef6", "#d7b5d8", "#df65b0", "#ce1256", "#fc9272", "#ef3b2c", 
    "#800026", "#fff7bc", "#fec44f", "#d95f0e", "#a63603", "#74c476", "#238b45", 
    "#00441b", "#7fcdbb", "#084081", "#c6dbef", "#2b8cbe", "#016c59", "#bcbddc", 
    "#807dba", "#54278f", "#bdbdbd", "black")

tax_names <- as.character(unique(tmp$Supergroup))
names(colors_tax) <- tax_names
```


```{r make bar plot back, fig.width=12, fig.height=8}
# View(asv_GR)
by_supergroup_back <- asv_GR %>% 
  filter(Domain == "Eukaryota") %>%
  filter(Sample_or_Control != "Control") %>% 
  # unite(SampleIdentifier, VENT, SAMPLETYPE, sep = " ", remove = FALSE) %>% 
  mutate(Supergroup = ifelse(is.na(Supergroup), "Unknown Eukaryota", Supergroup),
         Phylum = ifelse(is.na(Phylum), "Unknown", Phylum),
         Phylum = ifelse(Phylum == "Alveolata_X", "Ellobiopsidae", Phylum),
         Supergroup = ifelse(Supergroup == "Alveolata", paste(Supergroup, Phylum, sep = "-"), Supergroup)) %>% 
  group_by(FeatureID, Taxon, Domain, Supergroup, Phylum, Class, Order, Family, Genus, Species,
           VENT, SITE, SAMPLETYPE, YEAR, DATASET, Type) %>% 
    summarise(SEQ_AVG_REP = mean(value)) %>% 
  ungroup() %>% 
  group_by(SITE, SAMPLETYPE, VENT, Supergroup, Type) %>% 
    summarise(SEQ_SUM = sum(SEQ_AVG_REP)) %>% 
  unite(Sample_substrate, VENT, Type, sep = " ") %>% 
  filter(SITE == "GordaRidge") %>% 
  ggplot(aes(x = Sample_substrate, y = SEQ_SUM, fill = Supergroup)) +
    geom_bar(stat = "identity", position = "fill", color = "black", width = 0.6) +
    facet_grid(. ~ SITE + SAMPLETYPE, space = "free_x", scales = "free") +
  theme_linedraw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1),
        strip.background = element_blank(), strip.text = element_text(color = "black")) +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_manual(values = colors_tax) +
  labs(title = "Microcolonizers and background", x = "", y = "Relative sequence abundance")

by_supergroup_substrate <- asv_GR %>% 
  filter(Domain == "Eukaryota") %>%
  filter(Sample_or_Control != "Control") %>% 
  # unite(SampleIdentifier, VENT, SAMPLETYPE, sep = " ", remove = FALSE) %>% 
  mutate(Supergroup = ifelse(is.na(Supergroup), "Unknown Eukaryota", Supergroup),
         Phylum = ifelse(is.na(Phylum), "Unknown", Phylum),
         Phylum = ifelse(Phylum == "Alveolata_X", "Ellobiopsidae", Phylum),
         Supergroup = ifelse(Supergroup == "Alveolata", paste(Supergroup, Phylum, sep = "-"), Supergroup)) %>% 
  group_by(FeatureID, Taxon, Domain, Supergroup, Phylum, Class, Order, Family, Genus, Species,
           VENT, SITE, SAMPLETYPE, YEAR, DATASET, Type) %>% 
    summarise(SEQ_AVG_REP = mean(value)) %>% 
  ungroup() %>% 
  group_by(SITE, SAMPLETYPE, VENT, Supergroup, Type) %>% 
    summarise(SEQ_SUM = sum(SEQ_AVG_REP)) %>% 
  unite(Sample_substrate, VENT, Type, sep = " ") %>% 
  filter(SITE != "GordaRidge") %>% 
  ggplot(aes(x = Sample_substrate, y = SEQ_SUM, fill = Supergroup)) +
    geom_bar(stat = "identity", position = "fill", color = "black", width = 0.6) +
    facet_grid(. ~ SITE + SAMPLETYPE, space = "free_x", scales = "free") +
  theme_linedraw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1),
        strip.background = element_blank(), strip.text = element_text(color = "black")) +
  scale_y_continuous(expand = c(0,0)) +
  scale_fill_manual(values = colors_tax) +
  labs(title = "Microcolonizers and background", x = "", y = "Relative sequence abundance")


```

### Taxonomy plot of background fluid
```{r}
plotly::ggplotly(by_supergroup_back)
```
### Taxonomy plot of microcolonizers

```{r}
plotly::ggplotly(by_supergroup_substrate)
```

### Plot only metazoa
```{r plot metazoa only, fig.width=12, fig.height=8}
metazoa_substrate <- asv_GR %>% 
  filter(Domain == "Eukaryota") %>%
  filter(Sample_or_Control != "Control") %>% 
  mutate(Supergroup = ifelse(is.na(Supergroup), "Unknown Eukaryota", Supergroup),
         Phylum = ifelse(is.na(Phylum), "Unknown", Phylum),
         Phylum = ifelse(Phylum == "Alveolata_X", "Ellobiopsidae", Phylum),
         Supergroup = ifelse(Supergroup == "Alveolata", paste(Supergroup, Phylum, sep = "-"), Supergroup)) %>% 
  group_by(FeatureID, Taxon, Domain, Supergroup, Phylum, Class, Order, Family, Genus, Species,
           VENT, SITE, SAMPLETYPE, YEAR, DATASET, Type) %>% 
    summarise(SEQ_AVG_REP = mean(value)) %>% 
  # Opisthokonts only
  filter(Supergroup == "Opisthokonta") %>%
  unite(Phylum_Class, Phylum, Class, sep = "-") %>% 
  ungroup() %>% 
  group_by(SITE, SAMPLETYPE, VENT, Type, Supergroup, Phylum_Class) %>% 
    summarise(SEQ_SUM = sum(SEQ_AVG_REP)) %>% 
  unite(Sample_substrate, VENT, Type, sep = " ") %>% 
  ggplot(aes(x = Sample_substrate, y = SEQ_SUM, fill = Phylum_Class)) +
    geom_bar(stat = "identity", position = "fill", color = "black", width = 0.9) +
    facet_grid(. ~ SITE + SAMPLETYPE, scale = "free", space = "free") +
  theme_linedraw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1),
        strip.background = element_blank(), strip.text = element_text(color = "black")) +
  scale_y_continuous(expand = c(0,0)) +
  # scale_fill_manual(values = c("#f1eef6", "#d7b5d8", "#df65b0", "#ce1256", "#fc9272", "#ef3b2c", 
  #   "#800026", "#fff7bc", "#fec44f", "#d95f0e", "#a63603", "#74c476", "#238b45", 
  #   "#00441b", "#7fcdbb", "#084081", "#c6dbef", "#2b8cbe", "#016c59", "#bcbddc", 
  #   "#807dba", "#54278f", "#bdbdbd", "black")) +
  labs(title = "Average across replicates", x = "", y = "Relative sequence abundance")
```

```{r}
plotly::ggplotly(metazoa_substrate)
```

# What ASVs are shared across samples?
```{r}
library(ggupset)
```

```{r, fig.height=10, fig.width=13}
asv_GR %>% 
  filter(Domain == "Eukaryota") %>%
  filter(Sample_or_Control != "Control") %>% 
  mutate(Supergroup = ifelse(is.na(Supergroup), "Unknown Eukaryota", Supergroup),
         Phylum = ifelse(is.na(Phylum), "Unknown", Phylum),
         Phylum = ifelse(Phylum == "Alveolata_X", "Ellobiopsidae", Phylum),
         Supergroup = ifelse(Supergroup == "Alveolata", paste(Supergroup, Phylum, sep = "-"), Supergroup)) %>% 
  group_by(FeatureID, Supergroup, 
           VENT, Type) %>% 
    summarise(SUM = sum(value)) %>% 
  ungroup() %>%
  unite(Sample_substrate, VENT, Type, sep = " ") %>% 
  distinct(FeatureID, Supergroup, SUM, Sample_substrate, .keep_all = TRUE) %>% 
  group_by(FeatureID, Supergroup) %>% 
  summarise(Sample_substrate = list(Sample_substrate)) %>% 
  ggplot(aes(x = Sample_substrate)) +
    geom_bar(color = "black", width = 0.8, aes(fill = Supergroup)) +
    scale_x_upset(n_intersections = 30)
```

```{r, fig.height=10, fig.width=13}
# unique(asv_GR$VENT)
asv_GR %>% 
  filter(Domain == "Eukaryota") %>%
  filter(!(Type == "")) %>% 
  filter(Sample_or_Control != "Control") %>% 
  mutate(Supergroup = ifelse(is.na(Supergroup), "Unknown Eukaryota", Supergroup),
         Phylum = ifelse(is.na(Phylum), "Unknown", Phylum),
         Phylum = ifelse(Phylum == "Alveolata_X", "Ellobiopsidae", Phylum),
         Supergroup = ifelse(Supergroup == "Alveolata", paste(Supergroup, Phylum, sep = "-"), Supergroup)) %>% 
  group_by(FeatureID, Supergroup, 
           VENT, Type) %>% 
    summarise(SUM = sum(value)) %>% 
  ungroup() %>%
  unite(Sample_substrate, VENT, Type, sep = " ") %>% 
  distinct(FeatureID, Supergroup, SUM, Sample_substrate, .keep_all = TRUE) %>% 
  group_by(FeatureID, Supergroup) %>% 
  summarise(Sample_substrate = list(Sample_substrate)) %>% 
  ggplot(aes(x = Sample_substrate)) +
    geom_bar(color = "black", width = 0.8, aes(fill = Supergroup)) +
    scale_x_upset(n_intersections = 30)
```

# Import temperature data and other information for each microcolonizer
ibuttons deployed with each microcolonizer.

```{r}
sample_list <- read.delim("input-data/MC-samplelist.txt")
sample_metadata <- read.delim("input-data/MC-collect-info.txt")
amy <- c("Quartz1", "Basalt", "Olivine", "Pyrite")
sample_list_complete <- sample_list %>% 
  left_join(sample_metadata) %>% 
  separate(Sample_type, c("Sample_material", "Sample_type"), sep = "-") %>% 
  mutate(sample_purpose = case_when(
    (Sample_material %in% amy & Sample_type == "DNA") ~ "16S",
    (!(Sample_material %in% amy) & Sample_type == "DNA") ~ "18S",
    (Sample_material %in% amy & Sample_type == "SEM") ~ "Prok SEM",
    (!(Sample_material %in% amy) & Sample_type == "SEM") ~ "Euk SEM"
  ))
# write_delim(sample_list_complete, path = "sample-list-MC.txt", delim = "\t")
```
## Process ibutton data
```{r}
list_mc <- read.delim("input-data/microcolonizer_list.txt")
head(list_mc)
```
Deployment of all Microcolonizers was at 2019-05-29, 20:59:57.659 (UTC) (also 8:59 PM UTC), which is 4:59PM EST or 16:59 PM EST

Function to import all ibutton data raw and process.
```{r}
# Recovered microcolonizers by ibutton IDs:
## 1, 2, 3, 4, 5, 6
recovered <- c("2019-06-05 14:19:00","2019-06-04 16:04:00","2019-06-04 16:11:00","2019-06-05 14:23:00","2019-06-06 18:28:00","2019-06-06 18:31:00")
# Sys.timezone()
# tmp_2 <- logger2 %>% add_column(MC = 2)
# 
# logger1 <- read.csv("input-data/Logger1Data.csv", skip = 19)
# logger2 <- read.csv("input-data/Logger2Data.csv", skip = 19)
# x <- 1
mc_ids <- c(1, 2, 3, 4, 5, 6)

for(num in mc_ids){
  log_file <- read.csv(paste("input-data/Logger", num, "Data.csv", sep = ""), skip = 19)  
  cat("Reading in log file number", num, "\n")
  log_out <- log_file %>% 
    add_column(MC = num) %>% 
    mutate(Parsed_time_EST = parse_date_time(Date.Time, "%m/%d/%y %H:%M:%S p", tz = "America/New_York")) %>%
    # Filter out irrelevant date before deployment
    filter(Parsed_time_EST > "2019-05-29 18:59:00") %>% 
    filter(
      Parsed_time_EST < recovered[[num]]
    )
  if(!exists("log_files_all")){
    log_files_all <- log_out
  } else{
    log_files_all <- bind_rows(log_files_all, log_out)
  }
}
# rm(log_out); rm(log_files_all)
# head(log_out)
# View(log_files_all)


# Factor by colors and pairs of MCs
log_files_all$MC_ORDER <- factor(log_files_all$MC, levels = mc_ids)
mc_col <- c("#d7301f", "#4a1486", "#9e9ac8", "#fc8d59", "#2171b5", "#6baed6")
names(mc_col) <- mc_ids

```

Graph microcolonizer temperatures.
```{r}
ggplot(log_files_all, aes(x = Parsed_time_EST, y = Value, group = as.factor(MC_ORDER), color = as.factor(MC_ORDER))) +
  geom_path() +
  scale_color_manual(values = mc_col) +
  theme_classic(base_size = 14) +
  labs(x = "", y = "Temperature •C") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        legend.title = element_blank())
```

## Temperature data
```{r}
ggplot(log_files_all, aes(x = Parsed_time_EST, y = Value, color = as.factor(MC_ORDER))) +
  geom_step() +
  scale_color_manual(values = mc_col) +
  theme_classic(base_size = 14) +
  labs(x = "", y = "Temperature •C") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        legend.title = element_blank())
# ?geom_smooth
```








# Session end
```{r}
sessionInfo()
```

